---
title: "01_DataAquisition"
output: html_document
date: "2025-04-28"
---

Purpose: To filter the data and put together the files for imputation by funspace

Inputs: 
RData/Input_Files.RData - from Dietzel data 
coral_popsize.xlsx - from Dietzel supplementary data 
ctdb_1.1.1_data.csv - downloaded data from coral traits database 
coral_traits_raw.csv - recent 2025 download of the coral traits 

Outputs:
funspacedata_v2.csv - file with the mean traits of the corals with data in coral traits database matched to the support data from Dietzel


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(dplyr)
library(tidyverse)
```

```{r}
#load("../R_code_litreview_sem12024/R code Dietzel et al/RData/Input_Files.RData")

### Support Data Saved 
coral <- read_csv("../R_code_litreview_sem12024/coraldata.csv")
```

```{r}
file_path <- "../R_code_litreview_sem12024/coral_popsize.xlsx"

# Read the names of all sheets in the Excel file
sheet_names <- excel_sheets(file_path)
print(sheet_names)  # Print the names of the sheets

# Read each sheet into a list of data frames
data_list <- lapply(sheet_names, function(sheet) {
  read_excel(file_path, sheet = sheet)
})

# Optionally, name the list elements after the sheet names
names(data_list) <- sheet_names

# Access individual data frames by sheet name
species_abundance <- data_list[["Species abundances"]]
population_sizes <- data_list[["Population sizes"]]
habitat_maps <- data_list[["Reef habitat maps"]]

coraldata <- merge(coral, population_sizes, by = "Species")

coraldata <- coraldata%>%
  dplyr::select(Species, GrowthForm_CTD, Abund_CTD, MeanPCcover, family_molecules, family_morphology, larval)
  
```

coraltraits data 
```{r}
coraltraits_download <- read_csv("../methods/ctdb_1.1.1_data.csv")

coraltraits_download <- coraltraits_download %>%
  rename(Species = specie_name)

coraltraits_raw <- read_csv("../R_code_litreview_sem12024/coral_traits_raw.csv")

coraltraits_raw <- coraltraits_raw %>%
  rename(Species = specie_name)

##Making conversions for standard units 
relevant_traits <- c("Growth rate", "Colony fecundity", "Colony area", 
                     "Polyps per area", "Polyp fecundity", "Eggs per area", 
                     "Colony maximum diameter", "Corallite width minimum", 
                     "Corallite width maximum")

filtered_data <- coraltraits_raw %>%
  filter(trait_name %in% relevant_traits)

filtered_data_hardcoral <- subset(filtered_data, subclass != "Octocorallia")

unit_summary <- filtered_data_hardcoral %>%
  group_by(trait_name) %>%
  summarise(units = paste(unique(standard_unit), collapse = ", "))

combined_traits <- bind_rows(filtered_data_hardcoral, coraltraits_download) %>%
  distinct()

combined_traits <- combined_traits%>%
  dplyr::select(Species, trait_name, value, observation_id)%>%
  distinct()%>%
  dplyr::select(-observation_id)

combined_traits$Numeric_Value <- as.numeric(as.character(combined_traits$value))

combined_traits_num <- combined_traits%>%
  filter(!is.na(Numeric_Value))%>%
  dplyr::select(-value)

mean_traits <- combined_traits_num%>%
  group_by(Species, trait_name)%>%
  mutate(Meanvalue = mean(Numeric_Value))%>%
  dplyr::select(-Numeric_Value)%>%
  distinct()

data_wide <- mean_traits %>%
  pivot_wider(names_from = trait_name, values_from = Meanvalue)

names(data_wide) <- gsub(" ", "_", names(data_wide))

data_wide <- data_wide%>%
  dplyr::select("Species", "Growth_rate", "Colony_fecundity","Colony_area", "Polyps_per_area", "Polyp_fecundity", "Eggs_per_area","Colony_maximum_diameter","Corallite_width_minimum", "Corallite_width_maximum")

funspace_data <- left_join(data_wide, coraldata, by = "Species")

write_csv(funspace_data, "../methods/funspacedata_v2.csv")
```

```
polyp <-data_wide%>%
  filter(!is.na(Polyp_fecundity))

area <-data_wide%>%
  filter(!is.na(Colony_area))

# Combine the three datasets
combined_data <- bind_rows(need, polyp, area)

# Remove duplicate rows
combined_data_unique <- combined_data %>%
  distinct()

# View result
combined_data_unique


write_csv(combined_data_unique, "../methods/funspacedata_v2.csv")
```

Determining number of entries for each trait 
```{r}
n_traits_combined <- filtered_data_hardcoral %>%
  dplyr::select(Species, trait_name, value) %>%
  group_by(trait_name) %>%
  summarise(
    n_entries = n(),                         # Total number of entries
    n_unique_species = n_distinct(Species)   # Unique species per trait
  )%>%
  filter(trait_name %in% c("Colony area", "Colony fecundity", "Polyp fecundity", "Polyps per area", "Eggs per area"))

n_traits_combined <- filtered_data %>%
  dplyr::select(Species, trait_name, value) %>%
  group_by(trait_name) %>%
  summarise(
    n_entries = n(),                         # Total number of entries
    n_unique_species = n_distinct(Species)   # Unique species per trait
  )%>%
  filter(trait_name %in% c("Colony area", "Colony fecundity", "Polyp fecundity", "Polyps per area", "Eggs per area"))
```
